An Analysis of Wind Energy Potential in the North Sea

### 2024-01-29 paper/paper.tex
one fix in the tex
reformat paper source

\iffalse
\section{Theoretical Background}

\subsection{Methods to Estimate the Parameters of the Weibull Distribution}

For estimating the two parameters of the Weibull distribution there are several methods available \citep{compestimators},  
of which we focus on three: The \emph{graphical method}, the \emph{maximum likelihood method} and the \emph{power density method}, also called \emph{energy pattern method}. \\
Let $(v_i)_{i \in \{1, \dots, n\}}$ be the family of observed wind speeds. \\
The graphical method (GM) exploits the fact that, after some transformations, 
equation \ref{eq:cdf} can be considered a linear equation: 
\begin{align}
    \underbrace{\ln ( - \ln (1- F(v))}_{:= y}= \beta \underbrace{\ln(v)}_{:=x}- \beta \ln(\lambda)
\end{align}
Then for all observed values $v_i$, the empirical CDF can be computed, 
such that we obtain the values $x_i$ and $y_i$, on which linear regression can be performed.  
Historically, the graphical method was convenient, because the regression line could be
determined by hand, avoiding heavy computation \citep{graphical}. \\

The maximum likelihood method (MLM) is a general method to estimate the parameters of any distribution, provided there are i.i.d. samples available.
This theoretically poses a hindrance to application in our case, as the wind speed measurements are clearly not i.i.d. . However, the method is still recommended and yields good results \citep{compestimators, graphical}.\\
Application of the maximum likelihood method to the Weibull distribution yields the equations
\begin{align}
    \frac{1}{\beta}=
    \left( \frac{\sum_{i=1}^n v_i^\beta\ln(v_i)}{\sum_{i=1}^n v_i^\beta}
    - \frac{\sum_{i=1}^n \ln(v_i)}{n}
    \right)
    \label{eq: beta}
\end{align}
and 
\begin{align}
    \lambda = \left( \frac{\sum_{i=1}^n v_i^\beta}{n}  \right)^{\frac{1}{\beta}}.
    \label{eq: lambda}
\end{align}
Then $\beta$ is chosen as the numerical solution to \ref{eq: beta}
(as there could be several solutions to \ref{eq: beta}, experience shows that the ones near $\beta=2$ are preferable \citep{newestimator})
and $\lambda$ is subsequently computed from \ref{eq: lambda}. \\

For the power density method (PDM) the mean and the third moment are used to derive 
the energy pattern factor
\begin{align}
     E_{pf}:= \frac{E[X^3]}{E[X]^3}=\frac{ \Gamma \left( 1 + \frac{3}{\beta} \right)}{ \Gamma \left( 1 + \frac{1}{\beta} \right)^3}.
\end{align}
The mean and third moment can be estimated 
by $\bar v= \sum_{i=1}^n v_i$ and $\bar{v^3}= \sum_{i=1}^n  v_i^3$ respectively. 
Then $\beta$ could be found numerically; However, \citep{newestimator} 
proposes 
\begin{align}
    \beta= 1 + \frac{3.69}{E_{pf}^2}
\end{align}
as an approximate solution, and thereby introduced the power density method. 
Subsequently, $\lambda$ is the solution to $\bar v \approx E[X]=\Gamma \left( 1 + \frac{1}{\beta} \right)$.
The advantage of this method lies in its simplicity, as it does not require an iterative procedure.

\subsubsection{Goodness-of-Fit Criteria}
To evaluate the goodness-of-fit of the Weibull distribution to the wind speed frequencies, 
we first encounter the problem that we only have a finite number of samples of wind speeds, 
whereas the Weibull PDF is a continuous function.
This issue can be resolved by binning the samples and thus obtaining a categorical distribution. 
Then, for each bin, the empirical probability $p_i$ can be compared to the probability $\hat{p_i}$ obtained by the model. 
To obtain an appropriate number of bins, we apply the square root rule.\\
As a criterion for goodness-of-fit, the literature
\citep{review, compestimators,newestimator} often uses the root-mean-square error (RMSE),
\begin{align}
    \mathrm{RMSE}=\sqrt{ \frac{1}{k}\sum_{i=1}^k(p_i-\hat p_i)^2}
\end{align}
treating the problem as that of an ordinary regression.
However, as the issue at hand is the fitting of a probability distribution, the Kullback-Leibler Divergence,
comparing the ground-truth distribution $P$ to the estimated distribution $\hat P$,
    \begin{align}
        \mathrm{D_{KL}}(P, \hat P)=\sum_{i=1}^k \hat p_i \cdot \log\left(\frac{p_i}{\hat p_i}\right)
    \end{align}
is especially suitable.\\
\textcolor{magenta}{need citations for this?}

\fi

@@

We determine the parameters $\lambda$ and $\beta$ for every month and year using 
%\textcolor{magenta}{suitable estimator and assess the goodness of fit}. \\
the maximum likelihood method applied to \eqref{eq:pdf} \cite{mohammadi2016assessing}.
Due to the seasonal variability of wind speed, the long-term trend analysis is conducted on the yearly 
(rather than monthly) estimated wind speed PDFs. For this, we employ a least-squares linear regression and assess the statistical significance through a two-sided permutation test at the $5\%$ level. 
Since the two Weibull parameters fully characterize the distribution, we are then able to use this analysis to infer trends in other quantities of interest, like mean and standard deviation.

@@

\iffalse
\begin{figure*}
    \subfigure[The linear regression of the yearly Weibull parameters shows a noticeable decrease in 
 \textcolor{TUred}{$\lambda$}, but less so in \textcolor{TUdarkblue}{$\beta$}.]
        {\includegraphics[width=0.475\textwidth]{fig/linreg_parameters.pdf}
        \label{fig:linreg}
    }
    \hfill
  \subfigure[The estimated Weibull PDF for each year and the fit to the empirical PDF in 2018]{
    \includegraphics[width=0.475\textwidth]{fig/yearly_weibull.pdf}
    \label{fig:yearly_weibull}
    }
  \caption{The yearly Weibull parameters in \ref{fig:linreg} correspond to the PDFs in \ref{fig:yearly_weibull}.}
\end{figure*}
\fi

\iffalse

\subsection{Weibull Parameter Estimation}

The goals of this subsection are two-fold: Firstly to assess the overall fit of the Weibull distribution to the observed wind speed frequencies, 
and secondly, which of the estimators presented in 
\textcolor{magenta}{\ref{subsubsec: estimators}}
is most suitable. 
Since the models for the wind speed distribution are set to be used for 
monthly and yearly time periods, the goodness-of-fit is also evaluated at monthly and yearly basis.
Then the mean of the respective RMSE values and KL-Divergence values are used as the metric of comparison. 
The results are depicted in Figure \ref{tab:metrics}.

\begin{table}
    \begin{center}
        \begin{tabular}{l || c|c|c}
           Criterion & GM & MLM & PDM \\
            \hline
             av. yearly RMSE & 0.0020 & 0.0020 & 0.0020 \\
             av. yearly KL-Divergence & 0.0499 & 0.0455 & 0.0456 \\
             av. monthly RMSE & 0.0057 & 0.0050 & 0.0053 \\
             av. yearly KL-Divergence & 0.0544 & 0.0436  & 0.0462
        \end{tabular}
    \end{center}
    \caption{The average goodness-of-fit is best for the maximum likelihood method (MLM), closely followed by the probability density method (PDM) and 
    worst for the graphical method (GM). }\label{tab:metrics}
\end{table}

To put the values of the RMSE into perspective: As the number of bins are chosen according to the square root rule, the expected probability 
of a bin for the yearly time frame is approximately $\frac{1}{[\sqrt{6 \cdot 24 \cdot 365 }]} \approx 0.44 \%$ and for the monthly time frame it is 
 $\frac{1}{[\sqrt{6 \cdot 24 \cdot 30 }]} \approx 1.52 \%$ .

We can conclude that the maximum likelihood method provides the best results, 
while the power density method performs only slightly worse. 
Hence, we chose the maximum likelihood estimator for use in further analyses.

\fi

@@

For the available years (1999-2018) we estimated the parameters of the Weibull distribution on a monthly and yearly basis. 
\begin{figure}
    \centering 
    \includegraphics{fig/monthly_weibull.pdf}
    \caption{The monthly estimated Weibull distributions in the year 2010. The parameter $\lambda$ decreases in summer, whilst $\beta$ decreases, resulting in a shift to the left and lower standard deviation.}
    \label{fig:monthly_weibull}
\end{figure}
Figure \ref{fig:monthly_weibull} depicts the resulting PDFs for the months of 2010 as well 
as the empirical histogram in December.
When applying the square root rule to the number of bins,
 this fit has an expected relative error per bin of $18 \%$. 
 Averaged over the monthly and yearly distributions,
 the expected relative error per bin is  $22 \%$ and $23 \%$, respectively.

@@

\subsection{Gaussian Process Regression}
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=0.95\textwidth]{fig/gp_pred.pdf}
    \caption{The \textcolor{TUred}{mean} and \textcolor{TUred}{standard deviation} of the posterior distribution $\mathcal{N}(f(x); m', k')$ evaluated at $x$ is plotted together with the linearly interpolated \textcolor{TUdarkblue}{ground truth}. Starting at the \textcolor{TUdarkgreen}{data cut-off}, the model is extrapolating (predicting) the power density into the future.}
    \label{fig:gp-pred}
\end{figure*}

%For assessing the viability of wind turbine (parks), one might also be interested in the weekly or daily values  and, even more importantly, in a forecast. Hence, for interpolating the daily expected power density, we employ a Gaussian Process Regression model to perform a time-series regression. In addition to interpolating, we also use the same model for extrapolation and thereby making predictions about the future daily expected power density.
The first step for conducting a GP regression is to decide on a kernel $k$ for the prior \eqref{eq:gp_prior}. By analysing the plotted raw monthly expected power density, we identify a yearly periodic trend and a slight global downward trend. The former can be captured by the Exponential-Sine-Squared kernel (ES) \cite{MacKay1998IntroductionTG}. Supported by the empirical data, we choose a periodicity of $365 \, \text{days}$ and an amplitude of $450 \, \frac{\textrm{W}}{\text{m}^2}$. For the length scale, we settle for a rather small value ($< 2 \, $ days) in order to capture local patterns in the periodicity more accurately. For allowing the amplitude to slowly decay over long periods of time, we multiply the ES kernel by a very smooth (length scale $> 50 \, \text{years}$) Squared-Exponential (SE) kernel. For reproducing the global downwards trend observed in section \ref{sec:trend}, we also choose a very smooth Squared-Exponential kernel with an output scale of $100 \, \frac{\textrm{W}}{\text{m}^2}$. For accounting for the monthly midterm trend, we add a Rational-Quadratic kernel \cite{rasmussen-williams-gp} with a scale of $10 \, \frac{\textrm{W}}{\text{m}^2}$ and a length scale of one month. Finally, the additive kernel, consisting of a White-Noise (WN) and a Matern kernel (MA) \cite{abramowitz1968handbook}, is responsible for explaining measurement noise in the data and thus also prevents overfitting. Thereby, the kernel has the form
$$k = \underbrace{\text{ES} \cdot \text{SE}}_{\text{yearly trend}} + \underbrace{\text{RQ}}_\text{monthly trend} + \underbrace{\text{ES}}_\text{longterm trend} + \underbrace{\text{WN} + \text{MA}}_{\text{noise}}.$$
By drawing samples\footnote{with the mean of the data added.} $f_i \sim \mathcal{GP}(f; 0, k)$ from the prior distribution \eqref{eq:gp_prior}, we gauge how accurately our kernel choice fits to the given data. The results are shown in Figure \ref{fig:gp-samples}. We can observe that the amplitude and periodicity of some of the samples mirrors the data's in an accurate manner, supporting our choice of hyperparameters. After conditioning the Gaussian process prior on the normalised data, we obtain the posterior distribution $\mathcal{GP}(f; m', k')$ \eqref{eq:gp_posterior}. The evaluated posterior mean and standard deviation can be found in Figure \ref{fig:gp-pred}. We can observe that the mean function is following the periodicity of the ground truth without overfitting extreme power densities. Furthermore, the uncertainty of the model (standard deviation) increases gradually the further the prediction is from the data. Notably, the model also captures the global downward trend. When comparing the prediction to the test data in the years 2017 and 2018, we can observe that the prediction fits the ground truth accurately and all data points are well within one standard deviation of the mean.

>>>>

@@

We determine the parameters $\lambda$ and $\beta$ for every month and year using the maximum
likelihood method applied to \eqref{eq:pdf} \cite{mohammadi2016assessing}.  Due to the seasonal
variability of wind speed, the long-term trend analysis is conducted on the yearly (rather than
monthly) estimated wind speed PDFs. For this, we employ a least-squares linear regression and assess
the statistical significance through a two-sided permutation test at the $5\%$ level.  Since the two
Weibull parameters fully characterize the distribution, we are then able to use this analysis to
infer trends in other quantities of interest, like mean and standard deviation.

@@

@@

For the available years (1999-2018) we estimated the parameters of the Weibull distribution on a monthly and yearly basis. 
\begin{figure}
    \centering 
    \includegraphics{fig/monthly_weibull.pdf}
    \caption{The monthly estimated Weibull distributions in the year 2010. The parameter $\lambda$ decreases in summer, whilst $\beta$ decreases, resulting in a shift to the left and lower standard deviation.}
    \label{fig:monthly_weibull}
\end{figure}
Figure \ref{fig:monthly_weibull} depicts the resulting PDFs for the months of 2010 as well as the
empirical histogram in December.  When applying the square root rule to the number of bins,
this fit has an expected relative error per bin of $18 \%$.  Averaged over the monthly and yearly
distributions, the expected relative error per bin is  $23 \%$ and $22 \%$, respectively.

@@

\subsection{Gaussian Process Regression}
\begin{figure*}[htb!]
    \centering
    \includegraphics[width=0.975\textwidth]{fig/gp_pred.pdf}
    \caption{The \textcolor{TUred}{mean} and \textcolor{TUred}{standard deviation} of the posterior distribution $\mathcal{N}(f(x); m', k')$ evaluated at $x$ is plotted together with the linearly interpolated \textcolor{TUdarkblue}{ground truth}. Starting at the \textcolor{TUdarkgreen}{data cut-off}, the model is extrapolating (predicting) the power density into the future.}
    \label{fig:gp-pred}
\end{figure*}

%For assessing the viability of wind turbine (parks), one might also be interested in the weekly or daily values  and, even more importantly, in a forecast. Hence, for interpolating the daily expected power density, we employ a Gaussian Process Regression model to perform a time-series regression. In addition to interpolating, we also use the same model for extrapolation and thereby making predictions about the future daily expected power density.
The first step for conducting a GP regression is to decide on a kernel $k$ for the prior
\eqref{eq:gp_prior}. By analysing the plotted raw monthly expected power density, we identify a
yearly periodic trend and a slight global downward trend. The former can be captured by the
Exponential-Sine-Squared kernel (ES) \cite{MacKay1998IntroductionTG}. Supported by the empirical
data, we choose a periodicity of $365 \, \text{days}$ and an amplitude of $350 \,
\frac{\textrm{W}}{\text{m}^2}$. For the length scale, we settle for a rather small value ($< 2 \, $
days) in order to capture local patterns in the periodicity more accurately. For allowing the
amplitude to slowly decay over long periods of time, we multiply the ES kernel by a very smooth
(length scale $> 50 \, \text{years}$) Squared-Exponential (SE) kernel. For reproducing the global
downwards trend observed in section \ref{sec:trend}, we also choose a very smooth
Squared-Exponential kernel with an output scale of $100 \, \frac{\textrm{W}}{\text{m}^2}$. For
accounting for the monthly midterm trend, we add a Rational-Quadratic kernel
\cite{rasmussen-williams-gp} with a scale of $10 \, \frac{\textrm{W}}{\text{m}^2}$ and a length
scale of one month. Finally, the additive kernel, consisting of a White-Noise (WN) and a Matern
kernel (MA) \cite{abramowitz1968handbook}, is responsible for explaining measurement noise in the
data and thus also prevents overfitting. Thereby, the kernel has the form 
$$k = \underbrace{\text{ES} \cdot \text{SE}}_{\text{yearly trend}} 
+ \underbrace{\text{RQ}}_\text{monthly trend} 
+ \underbrace{\text{ES}}_\text{longterm trend} 
+ \underbrace{\text{WN} 
+ \text{MA}}_{\text{noise}}.$$
By drawing samples\footnote{with the mean of the data added.} $f_i \sim
\mathcal{GP}(f; 0, k)$ from the prior distribution \eqref{eq:gp_prior}, we gauge how accurately our
kernel choice fits to the given data. The results are shown in Figure \ref{fig:gp-samples}. We can
observe that the amplitude and periodicity of some of the samples mirrors the data's in an accurate
manner, supporting our choice of hyperparameters. After conditioning the Gaussian process prior on
the normalised data, we obtain the posterior distribution $\mathcal{GP}(f; m', k')$
\eqref{eq:gp_posterior}. The evaluated posterior mean and standard deviation can be found in Figure
\ref{fig:gp-pred}. We can observe that the mean function is following the periodicity of the ground
truth without overfitting extreme power densities. Furthermore, the uncertainty of the model
(standard deviation) increases gradually the further the prediction is from the data. Notably, the
model also captures the global downward trend. When comparing the prediction to the test data in the
years 2017 and 2018, we can observe that the prediction fits the ground truth accurately and all
data points are well within one standard deviation of the mean.
